---
title: "Power analysis"
author: "PSA 008"
date: "R&R v. 5"

bibliography: references.bib  

output:
  html_document:
    number_sections: true
    toc: true
    code_folding: hide
---

# README {-}

Here, we take the output of the power simulations and generates the plots used to determine the sample size. For each research question, this corresponds to the appropriate number of both countries (K = 30, 40, 50) and participants per country (N = 100, 150, 200, 250).

```{r, packages, message = FALSE}
## packages
library(tidyverse)

## read csv files and save to object
for(rq in c("rq1", "rq2", "rq3")) {
    assign(paste("pwr_", rq, sep = ""), 
           read_csv(paste("./pwr-", rq, ".csv", sep = "")))
}

```

# Research Question 1

For RQ1, we considered the effect of minimal group (in-group vs out-group) on one of the three outcomes. We used a conservative minimal group effect size of d = 0.19, implying a beta for minimal group = 0.1. The effect size was based on that reported for experiments lacking mutual independence in @Balliet-et-al_2014 (d = 0.19), which is about half the effect size of bias in allocation tasks for artificial groups reported in @Lane_2016 (d = 0.37). As the minimal group effect was assumed to vary across countries, we allowed random slopes for the minimal group effect to vary within country (slope variance = 0.01, 0.02, 0.03). 

```{r, rq1-plot}
## add N = to facet label 
subjects_labeller <- function(x) paste0("N = ", x)

## plot
pwr_rq1 %>%
    mutate(countries = as.factor(countries)) %>%
    ggplot() +
    aes(x = beta_variance, y = mean, group = countries, colour = countries) +
    geom_linerange(aes(ymin = lower, ymax = upper)) +
    geom_point(size = 3) +
    geom_line(alpha = 0.5, size = 1.5) +
    geom_hline(yintercept = 0.95, linetype = 2) +
    geom_hline(yintercept = 0.90, linetype = 3) +
    scale_x_continuous(breaks = c(0.01, 0.02, 0.03)) +
    ylim(0.5, 1.00) +
    facet_wrap(. ~ subjects,
               labeller = as_labeller(subjects_labeller)) +
    scale_colour_brewer(palette = "Dark2",
                        name = "K",
                        na.translate = FALSE) +
    labs(x = "slope variance",
         y = "power") +
    labs(title = "RQ1 power analysis") +
    guides(colour = guide_legend(reverse = TRUE)) +
    theme_classic(
      ## base_size = 30 ## for legibility of printed plot, comment out size
    ) +
    theme(legend.position="bottom",
           panel.spacing = unit(0.5, "lines"),
          strip.text.x = element_text(margin = margin(unit(1, "lines"))),
          strip.background = element_blank(),
          strip.placement = "outside")

```

Figure 1 shows the power obtained for RQ1. We achieve power >90% for all combinations of effect size and slope variance when K ≥ 50. At N = 200, we achieve power >95% when the slope variance is ≤ 0.02 and K ≥ 40.

# Research Question 2

For RQ2, we considered the effect of a continuous moderator (e.g. self-esteem, one of six moderators we will analyse) on the minimal group effect. We conducted a sensitivity analysis, in which we varied the beta for the moderator (beta = 0.05, 0.1, 0.2) and the random slope for the moderator within country (slope variance = 0.01, 0.02, 0.03). The betas for the moderator were based on the observed correlation of .11 between self-esteem and the minimal group effect in our Pilot Study 2 data. 

```{r, rq2-plot}
## add N = to facet label 
subjects_labeller <- function(x) paste0("N = ", x)
beta_labeller <- function(x) paste0("beta = ", x)

## plot
pwr_rq2 %>%
    mutate(countries = as.factor(countries)) %>%
    ggplot() +
    aes(x = beta_variance, y = mean, group = countries, colour = countries) +
    geom_linerange(aes(ymin = lower, ymax = upper)) +
    geom_point(size = 3) +
    geom_line(alpha = 0.5, size = 1.5) +
    geom_hline(yintercept = 0.95, linetype = 2) +
    geom_hline(yintercept = 0.90, linetype = 3) +
    scale_x_continuous(breaks = c(0.01, 0.02, 0.03)) +
    # ylim(0.00, 1.00) +
    facet_wrap(beta ~ subjects,
               ncol = 4,
               labeller = labeller(subjects = as_labeller(subjects_labeller), beta = as_labeller(beta_labeller))) +
    scale_colour_brewer(palette = "Dark2",
                        name = "K",
                        na.translate = FALSE) +
    labs(x = "slope variance",
         y = "power") +
    labs(title = "RQ2 power analysis") +
    guides(colour = guide_legend(reverse = TRUE)) +
    theme_classic(
      ## base_size = 30 ## for legibility of printed plot, comment out size
    ) +
    theme(legend.position="bottom",
          axis.text.x = element_text(angle = 45, hjust = 1),
          panel.spacing = unit(0.5, "lines"),
          strip.text.x = element_text(margin = margin(unit(1, "lines"))),
          strip.background = element_blank(),
          strip.placement = "outside")
```

Figure 2 shows the power obtained for RQ2. We achieve power >90% for all combinations of effect size and slope variance when K ≥ 50 and beta ≥ 0.1. At N = 200, we achieve power >95% when the slope variance is ≤ 0.02 and K ≥ 40.


# References {-}

<div id="refs"></div>