---
title: "Analysis pipeline"
author: "PSA 008"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

This document describes the analysis pipeline for PSA 008 Minimal Groups. It does not include the power analysis, which is in a separate document. The purpose is to troubleshoot --- some errors might be an artifact of the artificial data. 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R package information

```{r load-packages, message = FALSE, warning = FALSE}
## list of packages required
packages <- c(
  "tidyverse", # data wrangling
  "lme4", # random effects models
  "lmerTest", # random effects models
  "metafor", # meta analysis
  "mixedpower", # estimating power in lme
  "simr" # simulating data
)
## create list of packages that have not been installed
new_packages <-
    packages[!(packages %in% installed.packages()[,"Package"])]
## install packages that have not been installed
if (length(new_packages)) 
    install.packages(new_packages, dependencies = TRUE)
## load all packages
sapply(packages, library, character.only = TRUE)

source("./custom-functions.R")

set.seed(1970) # to reproduce analysis

sessionInfo() # session info
``` 

## Input artificial data

We created an artificial dataset that contains the relevant variables in a separate script, which we can call here and in other files. Note the dataset here will include the variables used in the main analyses, not all collected variables. Furthermore, the dataset resembles the processed data, not the raw data from Qualtrics.

```{r, source-simulated-data}
## loads the simulated data, and wrangled dataframes for each RQ
source("./data-simulation.R")

## peek at data
head(fake_data)
```

The (minimal group) dependent measures are based on the
three dictator games (in-group--self, out-group--self,
in-group--out-group) and the average attitude towards in-group and
towards out-group. The resultant three (minimal group) measures are: difference
between in-group and out-group attitudes (att_bias), difference between
in-group–self and out-group–self decisions in the dictator game (dg_first_bias), and
the decision in the in-group–out-group dictator game (dg_third_bias).

## Research question 1

Let's run some summary statistics (of att_bias).
We then use outcome minimal bias (min_bias), which includes all three (standardised) measures of bias with the minimal groups. Measure can then be include as a random effect.

```{r, rq1-manipulation}
## peek at dataframe
head(df_rq1)

## distribution of one outcome by country
fake_data %>% 
  ggplot(aes(att_min_bias)) + 
  geom_density() +
  facet_wrap(. ~ country)

## descriptives of one outcome
fake_data %>% 
  group_by(country) %>% 
  summarise(mean = mean(att_min_bias, na.rm = TRUE), 
            sd = sd(att_min_bias, na.rm = TRUE)) %>% 
  mutate_if(is.numeric, ~round(., 2)) %>% 
  print(n = n_countries)
```

We can run the model either with measure combined or as dummy; I find it easier to extract country-level random effects if we use dummy.

```{r, rq1-combined-dv-group}
## model with country
model_country_all <- 

    lmerTest::lmer(min_bias ~ measure + 
                       (1 | id)  + 
                       (measure | country/lab),
                   data = df_rq1
                   ## , contrasts = list(measure = "contr.sum")
                   )

## summary
summary(model_country_all)

## dotplot
gg_caterpillar(ranef(model_country_all, condVar = TRUE), QQ = FALSE, 
              likeDotplot = FALSE)

## predict the scores based on the model
df_rq1 %>% 
  mutate(mdl = predict(model_country_all)) %>%
  ggplot(aes(min_bias, y = mdl, colour = country, group = country)) + 
  geom_smooth(se = F, method = lm) +
  theme_bw() +
  guides(colour = "none") +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank())
```

``` {r, rq1-combined-dv-dummy}
## country
model_country_dummy <- 

    lmerTest::lmer(min_bias ~ att_dummy + dg_first_dummy + 
                       (1 | id)  + 
                       (att_dummy + dg_first_dummy + dg_third_dummy | country/lab),
                   data = df_rq1)

## summary
summary(model_country_dummy)

## icc
## country variance / (residual + country variance)

## dotplot
gg_caterpillar(ranef(model_country_dummy, condVar = TRUE), QQ = FALSE, 
              likeDotplot = TRUE)

## graph with predicted country level min bias

## predict the scores based on the model
df_rq1 %>%
  mutate(mdl = predict(model_country_dummy)) %>%
  ggplot(aes(min_bias, y = mdl, colour = country, group = country)) + 
  geom_smooth(se = FALSE, method = lm) +
  theme_bw() +
  guides(colour = "none") +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank())

## graph of random effects with line of best fit

## save coefficients
coefs_model <- coef(model_country_dummy)

## print random effects and best line
## shown is just dg_third_dummy
coefs_model$country %>%
  mutate(country = rownames(coefs_model$country),
         intercept = `(Intercept)`) %>% 
  ggplot(aes(x = dg_third_dummy, y = intercept, label = country)) + 
  geom_point() + 
  geom_smooth(se = F, method = lm) +
  geom_label(nudge_y = 0.15, alpha = 0.5) +
  theme_bw()
```

The above will be repeated with covariates.

## Research question 2

### Individual level 

For the individual level analysis of research question 2, we have three main
moderators of interest: permeability, (in-group and out-group) trust, and
self-esteem. We compare various models:

```{r df-rq2}
## peek at dataframe
head(df_rq2)
``` 

```{r rq2-models}
## minimal model
model_rq2_min <-
    
    lmerTest::lmer(
        min_bias ~ measure +  
          (1 | id)  + 
          (self_esteem + trust + permeability | country/lab),
        data = df_rq2)

## model with esteem
model_rq2_esteem <-
    
    lmerTest::lmer(
        min_bias ~ measure * self_esteem +  
          (1 | id)  + 
          (self_esteem + trust + permeability | country/lab),
        data = df_rq2)

## model with trust
model_rq2_esteem_trust <-
    
    lmerTest::lmer(
        min_bias ~ measure * (self_esteem + trust) +  
          (1 | id)  + 
          (self_esteem + trust + permeability | country/lab),
        data = df_rq2)

## model permeability
model_rq2_permeability <-
    
    lmerTest::lmer(
        min_bias ~ measure * permeability +  
          (1 | id)  + 
          (self_esteem + trust + permeability | country/lab),
        data = df_rq2)

## maximal model
model_rq2_max <-
    
    lmerTest::lmer(
        min_bias ~ measure * (self_esteem + trust + permeability) +  
          (1 | id)  + 
          (self_esteem + trust + permeability | country/lab),
        data = df_rq2)

## compare models
anova(model_rq2_min, 
      model_rq2_esteem, 
      model_rq2_esteem_trust,
      model_rq2_permeability,
      model_rq2_max)

## excplicitly AIC comparison, although above considers it with ML
AIC(model_rq2_min, 
    model_rq2_esteem, 
    model_rq2_esteem_trust,
    model_rq2_permeability, 
    model_rq2_max)
```

``` {r, rq2-country-level-effects}
## select model and run the following

## graph with predicted country level min bias

## predict the scores based on the model
df_rq2 %>%
  mutate(mdl = predict(model_rq2_permeability)) %>%
  ggplot(aes(permeability, y = mdl, colour = country, group = country)) + 
  geom_smooth(se = FALSE, method = lm) +
  theme_bw() +
  guides(colour = "none") +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank())

## graph of random effects with line of best fit

## save coefficients
coefs_model <- coef(model_rq2_permeability)

## print random effects and best line
## shown is permeability
coefs_model$country %>%
  mutate(country = rownames(coefs_model$country),
         intercept = `(Intercept)`) %>% 
  ggplot(aes(x = permeability, y = intercept, label = country)) + 
  geom_point() + 
  geom_smooth(se = F, method = lm) +
  geom_label(nudge_y = 0.001, alpha = 0.5) +
  theme_bw()
``` 

The above will be repeated with covariates.

### Country level 

We see if country-level predictors (Hofstede's individualism, the strength of family ties, in-group--out-group trust, and the Kinship Intensity Indicator), are significantly correlated with the countries' MGE means measures. For example, with KII:

```{r, rq2-country}
## read in country level indicators
country_level_indicators <-
  
  read_csv("./country-level-indicators.csv")

## save coefficients
coefs_model <- coef(model_country_dummy)

## save intercepts
country_intercept <-
  
  coefs_model$country %>%
  mutate(country = rownames(coefs_model$country),
         intercept = `(Intercept)`)

## print random effects and best line
## shown is just dg_third_dummy
joined <- 
  
  left_join(country_intercept,
    country_level_indicators,
    by = "country")

## correlation
cor.test(joined$KII, joined$intercept)

## plot
joined %>%
  ggplot(aes(x = KII, y = intercept, label = country)) + 
  geom_point() + 
  geom_smooth(se = F, method = lm) +
  geom_label(nudge_y = 0.001, alpha = 0.5) +
  theme_bw()

```

## Research question 3

For research question 3, we assess whether real-world bias (towards
the nation and/or family) is predicted by minimal group bias. 

```{r measures-real-world-bias}
## peek at dataframe
head(df_rq3)
```

To avoid three-way interactions, we model each measure separately. We demonstrate the analysis with the attitude measures.

``` {r rq3}
## real-world
model_real_world <- 

    lmerTest::lmer(att_real_bias ~ att_min_bias * group_type + 
                       (1 | id)  + 
                       (att_min_bias + group_type | country/lab),
                   data = df_rq3)

## summary
summary(model_real_world)

## ## run anova
## anova(model_real_world, ddf = "Kenward-Roger")

``` 

```{r plot-rq3}

model_coefs <- 
  
  coef(model_real_world)$country %>% 
  rename(intercept = `(Intercept)`, slope = att_min_bias) %>% 
  rownames_to_column("country")

model_coefs

df_rq3_rand <- left_join(df_rq3, model_coefs, by = "country")

ggplot(data = df_rq3_rand, 
       mapping = aes(x = att_min_bias, 
                     y = att_real_bias, 
                     colour = lab)
       ) +
  geom_point(na.rm = T, alpha = 0.5) +
  geom_abline(aes(intercept = intercept, 
                  slope = slope,
                  colour = lab
                  ),
              size = 1.5
              ) +
  facet_wrap(. ~ group_type) +
  guides(colour = "none") +
  theme_bw()
```

The above will be repeated for minimal bias in both the first- and third-party
dictator games.