---
title: "Power analysis example"
author: "PSA 008"
date: "R&R v. 4"

output:
  html_document:
    number_sections: true
    toc: true
    code_folding: show
---

# README {-}

This file presents the code for the power analysis, for research question 1. The code uses one of the simulated datasets, namely one with 200 participants across 40 countries. We use this as an example as it represents a number of participants and countries we reasonably assume we can recruit from. Users are welcome to change the dataset (e.g. to 100 participants across 40 countries), as well as any of the parameters. In this section, we also include random slopes for fixed effects, to demonstrate how the inferences do not change significantly with their inclusion.

# Packages {-}

```{r, package-information, message = FALSE}
## R package information

## load-packages
## list of packages required
packages <- c(
  "tidyverse", # data wrangling
  "lme4", # random effects models
  "lmerTest", # random effects models
  "simr", # simulating data
  "pwr", # power test
  "furrr", # iterate over multiple inputs
  "progressr", # display progress bar
  "patchwork" # combine plots
)
## create list of packages that have not been installed
new_packages <-
    packages[!(packages %in% installed.packages()[,"Package"])]
## install packages that have not been installed
if (length(new_packages)) 
    install.packages(new_packages, dependencies = TRUE)
## load all packages
sapply(packages, library, character.only = TRUE)

source("./custom-functions.R") # load custom functions

set.seed(1970) # to reproduce analysis
```
The number of simulations is set to 10 for troubleshooting, but in order to reproduce the results of the power analysis used in the manuscript the number of simulations should be set to 1,000. This can be acheived by placing ## in front of the number of simulations you do not want.

```{r, n-simulations}
n_sim <- 10 # to troubleshoot
## n_sim <- 1000 # actual run
```

# Research question 1

This is an example of estimating power for RQ1 for a given dataset using a single set of parameters, with varying random slopes.

```{r, rq1, cache = TRUE, eval = FALSE}
## function to simulate data and run power analysis
simulation_swarm_rq1 <- function(n_ids = 200, n_countries = 40, beta = 0.19, random_intercept = 1.5, random_slope) {
      
    ## create subject ids
    ids <- (1:n_ids) # start with n subjects per country
    
    ## create list of country names
    country_names <-
        letters_beyond_single_digits(n_countries) %>%
        tibble() %>%
        pull()
    
    ## create dataframe
    fake_data <-
        tibble(expand.grid(id = ids, country = country_names)) %>%
        mutate(id = c(1:nrow(.)))
        
    ## number of participants
    n_total <- nrow(fake_data)
    
    ## RQ1 variables
    
    ## create one of the minimal group dependent measures, based on the
    ## average attitude towards in-group and towards out-group
    ## mean values are taken from pilot 02, sd set to 1
    
    ## add outcomes for RQ1
    fake_data <-
        fake_data %>%
        mutate(min_in = round(
                   rtruncnorm(n_total, mean = 6.74, sd = 1, min = 0, max = 20), 0),
               min_out = round(
                   rtruncnorm(n_total, mean = 5.83, sd = 1, min = 0, max = 20), 0)
               ) %>%
        mutate(min_bias = min_in - min_out)

    ## extract dataframe that has "n" participants across "c" countries
    ## make long-form dataframe
    df <-
        fake_data %>%
        pivot_longer(cols = c("min_in", "min_out"),
                     names_to = "group",
                     values_to = "decision") %>%
        mutate(group = if_else(group == "min_in", "in", "out")) %>%
        dplyr::select(id, country, group, decision)
    
    ## extract number of countries and participants per country
    ## to include in summary output
    n_country <- pull(count(unique(dplyr::select(df, country))))
    n_id <- pull(count(unique(dplyr::select(df, id))))/n_country

    ## sigma in models
    residual_sd <- 1
    
    ## fixed intercepts 
    fixed <-c(
        0 # intercept
      , beta # beta for group, based on standardized minimal group effect
    )
    
    ## random intercepts
    random <- list(
        random_intercept # for country
      , .40 # for id, estimated from pilot 02 data (= 0.389)
      , random_slope
      , 0 
    )
    
    ## construct model
    ## in final analyses, we will include (group | country):
    model_rq1 <-
        makeLmer(
            decision ~ group + (group | country) + (1 | id),
            fixef = fixed, 
            VarCorr = random, 
            sigma = residual_sd, 
            data = df                 
        )

    ## check overall power
    sim_rq1 <- 
        powerSim(
            model_rq1,
            test = fixed("group"),
            nsim = n_sim,
            alpha = 0.05/3
        )

    ## assign powersim to object for later use
    summary_sim_rq1 <-

    summary(sim_rq1) %>%
    mutate(n_ids = n_id,
           n_countries = n_country,
           beta_group = beta,
           random_country = random_intercept,
           random_slope = random_slope,
	   class = "rq1"
           )
    
}

## create tibble with all combinations of parameters
parameter_space_rq1 <-

    crossing(
        random_slope = c(-0.1, 0, 0.1, 0.5, 1.0) # for "group in country" random slope
    )

## run power analysis across all parameters
pwr_rq1 <-
     
    future_pmap_dfr(parameter_space_rq1,
                    simulation_swarm_rq1,
                    .options = furrr_options(seed = TRUE),
                    .progress = TRUE
                    )

## print output
print(pwr_rq1)

## save in one csv for later use
pwr_rq1 %>%
    write.csv(file = paste0("./pwr-rq1-random.csv"))
```

Here is the output for RQ1.

```{r, rq1-1000-results, echo = FALSE, message = FALSE,  results = "asis"}
## create plot
plot_power_rq1_random <-
  read_csv("pwr-rq1-random.csv") %>%  
  mutate(n_ids = as.factor(n_ids)) %>%
  ggplot() +
  aes(x = random_slope, y = mean, group = n_ids, colour = n_ids) +
  geom_point(size = 3) +
  geom_line(alpha = 0.5, size = 1.5) +
  geom_hline(yintercept = 0.95, linetype = 2) +
  ylim(0.00, 1.00) +
  scale_x_continuous(breaks=c(-0.1, 0, 0.1, 0.5, 1.0)) +
  scale_colour_brewer(palette = "Dark2",
                        name = "number of\nparticipants",
                        na.translate = FALSE) +
    labs(x = "random slope of group within country",
         y = "power (beta)") +
    guides(colour = guide_legend(reverse = TRUE)) +
    labs(title = "Power analysis RQ1 with random slopes") +
    theme_bw(
      ## base_size = 30 ## for legibility of printed plot, comment out size
    )
## print plot
plot_power_rq1_random

## save plot
## plot_power_rq1_random %>%
##     ggsave(filename = paste(Sys.Date(), "pwr-rq1-random.png", sep = "_"),  
##           bg = "transparent")
```
