---
title: "Power analysis"
author: "PSA 008"
date: "R&R v. 3"

bibliography: references.bib  

output:
  html_document:
    number_sections: true
    toc: true
    code_folding: show
---

# README {-}

*NOTE*: This is currently a forked version that YD is working on to explore the power space. These changes are mainly in the second section below, the Example Power Analysis Section. Any changes we keep will eventually need to be replicated in the full power analysis section. 

This document describes the power analysis for PSA 008 Minimal
Groups. It is split into three parts. 

Section 1 details the general approach, how various parameters were derived, and how the dataset was simulated.

Section 2 shows the code used for the power analysis output reported in the manuscript, as well as reporting the output of the power analysis. Specifically, the power analysis using the mixed-effects modelling approach. The code iterates over different combinations of number of participants, number of countries, effect sizes, and random intercepts. By default, the number of simulations is set to 1 so that this document can be compiled. To reproduce the power analysis results reported in the manuscript, this should be changed to 1,000. This is time and resource intensive, with each simulation taking approximately 5 hours to complete when using 1,000 simulations. We used the University of Oxford Advanced Research Computing (ARC) facility (http://dx.doi.org/10.5281/zenodo.22558). This is why users should troubleshoot using a smaller number of simulations, and can make use of the code in Section 3 to run examples.

Section 3 presents the code for the power analysis, for each research question. The code uses one of the simulated datasets, namely one with 200 participants across 40 countries. We use this as an example as it represents a number of participants and countries we reasonably assume we can recruit from. Users are welcome to change the dataset (e.g. to 100 participants across 40 countries), as well as any of the parameters such as the effect size and random intercept.

# Packages and session information {-}

```{r, package-information, message = FALSE, class.source = 'fold-hide'}
## R package information

## load-packages
## list of packages required
packages <- c(
  "tidyverse", # data wrangling
  "lme4", # random effects models
  "lmerTest", # random effects models
  "simr", # simulating data
  "pwr", # power test
  "patchwork" # combine plots
)
## create list of packages that have not been installed
new_packages <-
    packages[!(packages %in% installed.packages()[,"Package"])]
## install packages that have not been installed
if (length(new_packages)) 
    install.packages(new_packages, dependencies = TRUE)
## load all packages
sapply(packages, library, character.only = TRUE)

source("./custom-functions.R") # load custom functions

set.seed(1970) # to reproduce analysis

sessionInfo() # session info
```

# General information {#sec:parameters}

## Set parameters

As outlined in the analysis plan, we have three main research questions (RQs). RQ1 assesses whether intergroup bias with minimal groups (i.e. the minimal group effect) holds across the societies we will sample. RQ2 assesses whether the minimal group effect varies as a function of key moderators derived from the literature. RQ3 looks at the relationship between the minimal group effect and bias with real-world groups. Full details are in the manuscript.

For all research questions, the aim of the power analysis is to calculate the number of participants per country, given three parameters: significance level, power, and effect size.

* We set the significance level to $alpha = .05/3 = .017$. This is because there are three outcomes tested within each research question. In all cases, the outcome is a measure of bias, which is the difference between the in-group and the out-group in either first-party allocation, third-party allocation, or attitudes.

* Given our frequentist analysis plans, the a priori power is set to $beta = 0.95$ or higher.

* For RQ1 and RQ2, the outcome is the minimal group effect; we looked to the literature for an indication of effect size for the minimal group effect. We derived two values, a larger and smaller one. The larger value is based on @Balliet-et-al_2014, a meta-analysis of intergroup discrimination in behavioural economic games. The overall effect size of games involving experimentally created (i.e. minimal) groups was $d = 0.35 (0.27-0.42)$. The value is similar to the overall effect size of $d = 0.37 (0.28-0.45; n = 150)$ in games involving "artificial" (i.e. minimal) groups reported in @Lane_2016. We set the effect size to 0.16, which is approximately half the larger value and corresponds with the average effect size in experiments lacking mutual interdependence.

* For RQ3, the outcome is bias with real-world (i.e. national and family) groups. We derived an effect size based on literature reporting studies with national groups, as this is more common than with family groups. Effects sizes relevant to national groups in economic tasks are a meta-analysis and a recent cross-cultural study. @Lane_2016 also eports an overall effect size of "national" groups $d = 0.16 (0.04–0.29; n = 52)$, and @Romano-et-al_2021 conducted a study across 42 nations finding $d = 0.22 (0.19-0.25)$ (this compares in-group with out-group plus strangers; from supplementary figures, it appears that the difference between in-group and out-group is less than in-group and stranger, so the overall effect size for in-group vs. out-group is likely less than $0.22$). We are not aware of an overall effect size of family from a meta-analysis or large scale study to report an effect size. Thus, we set the effect size to the smaller value of 0.16, which is the same used for RQ1 and RQ2.

For all RQs, we used the "simr" package [@Green-McLeod_2016], which allows for the a simplified version of the planned statistical model to be specified and the power calculated for a varying number of participants per country, number of countries, and country intercept variances. For all these calculations, we used simulated data, described below.

The beta coefficients were also derived from the literature, where possible. For RQ1, the coefficient for the fixed effect of group was 0.3, based on a similar in-group vs. out-group comparison, namely the value of 0.29 reported for the in-group vs out-group/stranger comparison in Romano et al (2021). For RQ2, the beta coefficient for the fixed effect of self-esteem was 0.4. For this, we relied on a variable related to one of our predictors for which there was a reliable estimate, namely a variable related to trust. We estimated the effect of State Trust on the rate of cooperation using the meta-regression tool on the Cooperation Databank (https://app.cooperationdatabank.org/). The meta-regression (restricted ML estimator) reported a beta of 0.419 when conducted on 2022-11-22, which was composed of 11 effects and a total of 2,036 participants. For RQ3, the beta coefficients for the fixed effects of minimal group effect and of group type were both was 0.3, the same as for RQ1; their interaction was set to 1.0.
 
## Simulate data

The simulated data used the results from the second pilot to specify the outcome per country for RQ1 and RQ2 (i.e. minimal group bias) and for RQ3 (i.e. real-world bias). In addition, for RQ2 we generated one of the measures (self-esteem), for which we have a good cross-country estimate using a similar scale as reported in @Robins-et-al_2001.  

```{r, simulate-data}
## first generate the basic characteristics: id, country

## set the number of labs

## start with half value of number of interested collaborators at
## first submission which is n = 59
list_n_countries <- c(seq(from = 20, to = 60, by = 10))

##  vary number of participants per country
list_n_ids_per_country <- c(seq(from = 100, to = 400, by = 50))

## iteratively create datasets that vary in the number of countris
for (n_countries in list_n_countries) {

    ## and that vary in number of participants per country
    for (n_ids in list_n_ids_per_country) {
        
        ## create subject ids
        ids <- (1:n_ids) # start with n subjects per country
        
        ## create list of country names
        country_names <-
            
            letters_beyond_single_digits(n_countries) %>%
            tibble() %>%
            pull()
        
        ## create dataframe
        fake_data <-
            
            expand.grid(id = ids, country = country_names) %>%
            tibble() %>%
            mutate(id = c(1:nrow(.)))
        
        ## number of participants
        n_total <- nrow(fake_data)
        
        ## RQ1 variables
        
        ## create one of the minimal group dependent measures, based on the
        ## average attitude towards in-group and towards out-group
        
        ## mean and sd values are taken from pilot 02
        
        ## add outcomes for RQ1 and RQ2
        fake_data <-
            
            fake_data %>%
            mutate(min_in = 
                       round(
                           rtruncnorm(n_total, 
                                      mean = 6.74, sd = 4.25, min = 0, max = 20), 
                           0),
                   min_out = 
                       round(
                           rtruncnorm(n_total, 
                                      mean = 5.83, sd = 4.49, min = 0, max = 20), 
                           0)
                   ) %>%
            mutate(min_bias = min_in - min_out)
        
        ## RQ2 variables
        ## generate one of the measures, we use self-esteem as we have a good
        ## cross-country estimate using a similar scale
        
        ## generate self-esteem
        self_esteem_score <- 
            
            rnorm(n_ids
                , mean = 3.5, sd = 1 # adapted from Robins-et-al_2001
                  ) 
        
        ## add to artificial data frame
        ## each repeated n times for each country
        fake_data <-
            
            fake_data %>%
            mutate(self_esteem = rep(self_esteem_score, n_countries)
                   )
        ## RQ3 variables

        ## mean and sd values are taken from pilot 02
        
        ## add real-world measures
        fake_data <-
            
            fake_data %>%
            ## nation
            mutate(
                nat_in = 
                    round(
                        rtruncnorm(n_total, 
                                   mean = 6.29, sd = 4.26, min = 0, max = 20), 
                        0),
                nat_out = 
                    round(
                        rtruncnorm(n_total, 
                                   mean = 6.26, sd = 4.59, min = 0, max = 20), 
                        0)) %>%
            mutate(nat_bias = nat_in - nat_out) %>%
            ## family        
            mutate(
                fam_in = 
                    round(
                        rtruncnorm(n_total, 
                                   mean = 3, sd = 3, min = 0, max = 10), 
                        0),
                fam_out = 
                    round(
                        rtruncnorm(n_total, 
                                   mean = 1, sd = 3, min = 0, max = 10), 
                        0)) %>%
            mutate(fam_bias = fam_in - fam_out)
        
        ## assign to object for later use
        assign(
            ## name of object
            paste0("fake_data_n", n_ids, "_c", n_countries),
            ## object
            fake_data
        )
        
    }

}

## gather all fake datasets 
gather_dfs <- grep("fake_data_n", names(.GlobalEnv), value = TRUE)
list_dfs <- do.call("list", mget(gather_dfs))
```

# Complete power analysis

The first thing to set is the number of simulations. This is set to 1 for troubleshooting, but in order to reproduce the results of the power analysis used in the manuscript the number of simulations should be set to 1,000. This can be acheived by placing ## in front of the number of simulations you do not want.

```{r, n-simulations}
n_sim <- 1 # to troubleshoot
## n_sim <- 1000 # actual run
```

As described above, the outcome is set to the smaller of two effects, $d =$ 0.16.

```{r, outcomes}
## set effect sizes
es_outcome_larger <- 0.35
es_outcome_smaller <- 0.16
min_bias_es <- es_outcome_smaller
```

We conduct a power analysis for each research question (RQ), using a version of the pre-specified mixed-models we will conduct as part of the planned analyses. The specification was trimmed. Only the random intercept for country was included. No random slopes were included because of in the absence of reliable estimates from prior literature and computational issues. 

For each fixed effect we specify the beta coefficients, for each random effect the variance, and finally the error variance (\ie sigma). Coefficients for the effects of interest (group for RQ1, a single moderator for RQ2, the interaction of the minimal group effect and group type for RQ3) were based on expectations from the literature. The variance was set to multiple values. The error variance is set to 0.1 in all models.

We display the code that generates the output. Warning, this code below is computational intensive. It is advised that users run the code in Section 3, which is an example using a single set of parameters. We then print the output in a table for easy viewing. These pull in the `.csv` files saved in the repository, which contain the output from the the full run of 1,000 simulations.

## Codebook

Please refer to the following codebook for the headers in the power analysis output.

```{r, rq1-1000-codebook, echo = FALSE, message = FALSE, results = "asis"}
read_csv("pwr-rqs-codebook.csv") %>%
  knitr::kable(., caption = "Codebook for power analysis output tables")
```

## Research question 1

The output for RQ1 is generated by the following iterative code.

```{r, rq1-specs}
## sigma in models
residual_sd <- 1

## group beta
beta <- 0.3
## based on in-group vs out-group/stranger of beta 0.29 in Romano et
## al (2021) https://doi.org/10.1038/s41467-021-24787-1

## we set the lowest value for the variance to 0.05

## for sensitivity analysis
## create list of random intercepts
list_random_intercepts <- c(0.05, 0.30, 0.55, 0.80, 1.5)
```

```{r, rq1-iter, eval = FALSE}
## iterate over different random intercepts
for (random_intercept in list_random_intercepts) {
  
  ## iterate over different number of ids per country
  for (df in list_dfs) {
    
    ## dataframe
    ## extract dataframe that has "n" participants across "c" countries
    ## make long-form dataframe
    df <-
      
      df %>%
      pivot_longer(cols = c("min_in", "min_out"),
                   names_to = "group",
                   values_to = "decision") %>%
      mutate(group = if_else(group == "min_in", "in", "out")) %>%
      select(id, country, group, decision)
    
    ## fixed intercept 
    fixed <- c(
      min_bias_es # outcome
      , 0.4 # moderator
    )
    
    ## random intercepts
    random <- list(
      random_intercept  
    )
    
    ## construct model
    model_rq1 <-
      
      makeLmer(
        decision ~ group + (1 | country),
        fixef = fixed, 
        VarCorr = random, 
        sigma = residual_sd, 
        data = df                 
      )
    
    ## check power
    sim_rq1 <- 
      
      powerSim(
        model_rq1,
        test = fixed("group"),
        nsim = n_sim,
        alpha = 0.05/3
      )
    
    ## extract number of countries and participants per country
    n_country <- pull(count(unique(select(df, country))))
    n_id <- pull(count(unique(select(df, id))))/n_country
    
    ## assign powersim to object for later use
    summary_sim_rq1 <-
      
      summary(sim_rq1) %>%
      mutate(bias_es = min_bias_es,
             random_var = random_intercept,
             n_ids = n_id,
             n_countries = n_country,
             beta = beta,
             class = "rq1"
      )
    
    ## assign powercurve to object for later use
    assign(
      ## name of object
      paste0("pwr_rq1_es", min_bias_es,
             "_ri", random_intercept,
             "_n", n_id,
             "_c", n_country),
      ## summary
      summary_sim_rq1
    )
    
  }
}

## gather all power simulations 
gather_pwr_sim <- grep("pwr_rq1", names(.GlobalEnv), value = TRUE)
list_pwr_sim <- do.call("list", mget(gather_pwr_sim))

bind_rows(list_pwr_sim) %>%
    arrange(bias_es, random_var, n_ids, n_countries) %>%
mutate(powered = if_else(mean > 0.95, "yes", "no")) %>%
    select(n_ids, n_countries, bias_es,	beta, random_var,	trials, successes, mean,	lower, upper,	powered, class) %>%
    write_csv(path = "./pwr-rq1-sims-combined.csv")
```

Here is the output for RQ1. In sum, we are sufficiently powered across all parameters we have tested.

```{r, rq1-1000-results, echo = FALSE, message = FALSE, results = "asis"}
read_csv("pwr-rq1-sims-combined.csv") %>%
  arrange(n_ids) %>%
  mutate(powered = if_else(mean > 0.95, "yes", "no")) %>%
  select(n_ids, n_countries, bias_es, beta,	random,	trials, successes, mean, lower, upper, powered, class) %>%
  collapse_rows_df(n_ids) %>%
  knitr::kable(., caption = "RQ1 power analysis (n simulations = 1,000)")
```

## Research question 2

The output for RQ2 is generated by the following iterative code.

```{r, rq2-iter, eval = FALSE}
## sigma in models
residual_sd <- 1

## moderator beta
beta <- 0.4
## based on beta 0.419 from https://app.cooperationdatabank.org/

## this selected State Trust as independent variable, which as as of
## 2022-11-22, was composed of 11 effects and a total of 2036 participants

## we set the lowest value for the variance to 0.05

## for sensitivity analysis
## create list of random intercepts
list_random_intercepts <- c(0.05, 0.30, 0.55, 0.80, 1.5)

## iterate over different random intercepts
for (random_intercept in list_random_intercepts) {
  
  ## iterate over different number of ids per country
  for (df in list_dfs) {
    
    ## fixed intercept 
    fixed <- c(
      min_bias_es # outcome
      , 0.4 # moderator
    )
    
    ## random intercepts
    random <- list(
      random_intercept    
    )
    
    ## construct model
    model_rq2 <- 
      
      makeLmer(
        min_bias ~ self_esteem + (1 | country),
        fixef = fixed, 
        VarCorr = random, 
        sigma = residual_sd, 
        data = df                 
      )
    
    ## check power
    sim_rq2 <- 
      
      powerSim(
        model_rq2,
        test = fixed("self_esteem"),
        nsim = n_sim,
        alpha = 0.05/3
      )
    
    ## extract number of countries and participants per country
    n_country <- pull(count(unique(select(df, country))))
    n_id <- pull(count(unique(select(df, id))))/n_country
    
    ## assign powersim to object for later use
    summary_sim_rq2 <-
      
      summary(sim_rq2) %>%
      mutate(bias_es = min_bias_es,
             random = random_intercept,
             n_ids = n_id,
             n_countries = n_country,
             beta = beta,
             class = "rq2"
      )
    
    ## assign powercurve to object for later use
    assign(
      ## name of object
      paste0("pwr_rq2_es", min_bias_es,
             "_ri", random_intercept,
             "_n", n_id,
             "_c", n_country),
      ## summary
      summary_sim_rq2
    )
    
  }
  
}

## gather all power simulations 
gather_pwr_sim <- grep("pwr_rq2", names(.GlobalEnv), value = TRUE)
list_pwr_sim <- do.call("list", mget(gather_pwr_sim))

bind_rows(list_pwr_sim) %>%
  arrange(bias_es, random_var, n_ids, n_countries) %>%
mutate(powered = if_else(mean > 0.95, "yes", "no")) %>%
    select(n_ids, n_countries, bias_es,	beta, random_var,	trials, successes, mean,	lower, upper,	powered, class) %>%
    write_csv(path = "./pwr-rq2-sims-combined.csv")
```

Here is the output for RQ2. In sum, we are sufficiently powered across all parameters we have tested.

```{r, rq2-1000-results, echo = FALSE, message = FALSE, results = "asis"}
read_csv("pwr-rq2-sims-combined.csv") %>%
  arrange(n_ids) %>%
  mutate(powered = if_else(mean > 0.95, "yes", "no")) %>%
  select(n_ids, n_countries, bias_es, beta, random,	trials, successes, mean,	lower, upper,	powered, class) %>%
  collapse_rows_df(n_ids) %>%
  knitr::kable(., caption = "RQ2 power analysis (n simulations = 1,000)")
```

## Research question 3

Finally, the output for RQ3 is generated by the following iterative code.

```{r, rq3-iter, eval = FALSE}


## sigma in models
residual_sd <- 1

## effect size
real_bias_es <- es_outcome_smaller

## moderators beta
beta_mge <- 0.3
beta_group_type <- 0.3 # same as for rq1
beta_interaction <- 1.0

## we set the lowest value for the variance to 0.05 in our power
## analysis

## for sensitivity analysis
## create list of random intercepts
list_random_intercepts <- c(0.05, 0.30, 0.55, 0.80, 1.5)

## iterate over different random intercepts
for (random_intercept in list_random_intercepts) {
  

  
  ## iterate over different number of ids per country
  for (df in list_dfs) {
   
## adjust dataframe to contain relevant predictors 
## (MGE, group type)
df <-
  
  df %>%
  pivot_longer(cols = c("nat_bias", "fam_bias"),
               names_to = "group_type", values_to = "real_bias") %>%
  rename(MGE = min_bias) %>%
  separate(group_type, into = c("group_type", "bias")) %>%
  select(id, country, MGE, group_type, real_bias)

    ## fixed intercept 
    fixed <- c(
      real_bias_es # outcome
      , beta_mge # MGE
      , beta_group_type # group type
      , beta_interaction # interaction
    )
    
    ## random intercepts
    random <- list(
      random_intercept # for country
    )
    
    ## construct model
    model_rq3 <-
      
      makeLmer(
        real_bias ~ MGE * group_type + (1 | country), 
        # we use simplier random structure to make code easier to run
        fixef = fixed, 
        VarCorr = random, 
        sigma = residual_sd, 
        data = df
      )
    
    ## check power
    sim_rq3 <- 
      
      powerSim(
        model_rq3,
        test =  fcompare(~ MGE + group_type),
        nsim = n_sim,
        alpha = 0.05/3
      )
    
    ## extract number of countries and participants per country
    n_country <- pull(count(unique(select(df, country))))
    n_id <- pull(count(unique(select(df, id))))/n_country
    
    ## assign powersim to object for later use
    summary_sim_rq3 <-
      
      summary(sim_rq3) %>%
      mutate(bias_es = real_bias_es,
             random = random_intercept,
             n_ids = n_id,
             n_countries = n_country,
             class = "rq3"
      )
    
    ## assign powercurve to object for later use
    assign(
      ## name of object
      paste0("pwr_rq3_es", real_bias_es,
             "_ri", random_intercept,
             "_n", n_id,
             "_c", n_country),
      ## summary
      summary_sim_rq3
    )
    
  }
  
}

## gather all power simulations 
gather_pwr_sim <- grep("pwr_rq3", names(.GlobalEnv), value = TRUE)
list_pwr_sim <- do.call("list", mget(gather_pwr_sim))

bind_rows(list_pwr_sim) %>%
  arrange(bias_es, random_var, n_ids, n_countries) %>%
mutate(powered = if_else(mean > 0.95, "yes", "no")) %>%
    select(n_ids, n_countries, bias_es, random_var,	trials, successes, mean,	lower, upper,	powered, class) %>%
    write_csv(path = "./pwr-rq3-sims-combined-trial.csv")
```

Here is the output for RQ3. In sum, we are sufficiently powered across all parameters we have tested.

```{r, rq3-1000-results, echo = FALSE, message = FALSE,  results = "asis"}
read_csv("pwr-rq3-sims-combined.csv") %>%
  arrange(n_ids) %>%
  mutate(powered = if_else(mean > 0.95, "yes", "no")) %>%
  select(n_ids, n_countries, bias_es,	random,	trials, successes, mean,	lower, upper,	powered, class) %>%
  collapse_rows_df(n_ids) %>%
  knitr::kable(., caption = "RQ3 power analysis (n simulations = 1,000)")
```

# Example power analysis

YD playing with code from here onward only, haven't updated the iterative code above except to correct a code error in which the data reshaping was happening outside instead of inside the iteration (the for loop).

## Research question 1

### Mixed-effects model approach

This is an example of estimating power for RQ1 for a given dataset using a single set of parameters.

```{r, rq1, cache = TRUE}
## sigma in models
residual_sd <- 1

## group beta
beta <- es_outcome_smaller
## based on standardized minimal group effect

## we set the lowest value for the variance to 0.05

## for sensitivity analysis
## create list of random intercepts
list_random_intercepts <- c(0.05, 0.30, 0.55, 0.80, 1.5)
## select value from list_random_intercepts for example
random_intercept <- c(0.30)

## fixed intercept 
fixed <- c(
    0 # intercept
  , beta # group beta
)

## random intercepts, .30 for country and .39 for id; latter estimated from pilot data
random <- list(.30,.39)

## dataframe
## extract dataframe that has "n" participants across "c" countries
## make long-form dataframe
df <-

    fake_data_n200_c40 %>%
    pivot_longer(cols = c("min_in", "min_out"),
                 names_to = "group",
                 values_to = "decision") %>%
    mutate(group = if_else(group == "min_in", "in", "out")) %>%
    select(id, country, group, decision)

## extract number of countries and participants per country
## required for producing output
n_country <- pull(count(unique(select(df, country))))
n_id <- pull(count(unique(select(df, id))))/n_country

## construct model
## in final analyses, we will include (group | country):
## decision ~ group + (1 | id) + (group | country)
## but we use a simplier specification for the power analysis
model_rq1 <-
    
    makeLmer(
        decision ~ group + (1 | country) + (1|id),
        fixef = fixed, 
        VarCorr = random, 
        sigma = residual_sd, 
        data = df                 
    )

## check overall power
sim_rq1 <- 
    
    powerSim(
        model_rq1,
        test = fixed("group"),
        nsim = n_sim,
        alpha = 0.05/3
    )

sim_rq1

## assign powersim to object for later use
summary_sim_rq1 <-
    
    summary(sim_rq1) %>%
    mutate(bias_es = min_bias_es,
           random = random_intercept,
           n_ids = n_id,
           n_countries = n_country,
           beta = beta
           )

## summarise
summary_sim_rq1
```

### Meta-analytical approach

Meta-analytical approach based on "meta_analysis_power.R" script from @Quintana-Tiebel_2018 and @Valentine-et-al_2010.

```{r, rq1-meta-analysis}
## overall effect size
effect_size <- es_outcome_smaller

## number of countries
## lower limit: 20 is a reasonable minimal number based on previous PSA projects
## upper limit: 60, as 59 countries expressed an interest at first submission
range_countries <- c(seq(20, 60, 10))

## number of participants per treatment
## corresponding to half total number of participants
range_group <- c(25, 50, 75, 100, 125, 150, 175, 200)

## calculate critical z
c_z <- qnorm(p = (0.05/3)/2, # adjusted alpha, halved for two-tailed
             lower.tail = FALSE)

## create tibbles with estimated power

## across range of heterogeneity = tau squared
for (heterogeneity in c(.33, 1, 3)) {
    
    ## across range of countries
    for (n_k in range_countries) {
        
        ## across range of number of participants per treatment
        for (n_per_group in range_group) {
            
            eq1 <- 
                ((n_per_group + n_per_group)/((n_per_group) * (n_per_group))) + 
                ((effect_size^2)/(2*(n_per_group + n_per_group)))
            
            eq2 <- heterogeneity*(eq1)
        
            eq3 <- eq2 + eq1
            
            eq4 <- eq3/n_k
            
            eq5 <- (effect_size/sqrt(eq4))
        
            power <- (1 - pnorm(c_z - eq5)) # two-tailed
            
            power

            ## create tibbles
            assign(
                paste0("tbl_", n_per_group, "_", n_k, "_", heterogeneity),
                tibble(n = n_per_group, k = n_k, pwr = power, hg = heterogeneity))
            
        }
    }
    
}

## gather all tibbles 
gather_tbls <- grep("tbl", names(.GlobalEnv), value=TRUE)
tbls_list <- do.call("list", mget(gather_tbls))

## create plot
## YD: does not run, column names issue
plot_power <-

    bind_rows(tbls_list) %>%
    mutate(n = n*2) %>% # double to get total sample size
    mutate(k = as.factor(k)) %>%
    ggplot() +
    aes(x = n, y = pwr, group = k, colour = k) +
    geom_point(size = 3) +
    geom_line(alpha = 0.5, size = 1.5) +
    geom_hline(yintercept = 0.95, linetype = 2) +
    facet_wrap(. ~ hg,
               labeller = labeller(hg = as_labeller(c("0.33" = "0.33 (low heterogeneity)",
                                                      "1" = "1.00 (mid heterogeneity)",
                                                      "3" = "3.00 (high heterogeneity)"
                                                      )))) +
    scale_colour_brewer(palette = "Dark2",
                        name = "number of\ncountries",
                        na.translate = FALSE) +
    labs(x = "sample size (total n)",
         y = "power (Beta)") +
    guides(colour = guide_legend(reverse = TRUE)) +
    theme_bw(
      ## base_size = 33 ## for legibility of printed plot, comment out size
      ) 

## visualise plot
plot_power

 ## save plot # comment out to save in folder 
 plot_power %>%
     ggsave(filename = paste(Sys.Date(), "power-rq1.png", sep = "_"),  
            bg = "transparent")
```

## Research question 2

This is an example of estimating power for RQ2 for a given dataset using a single set of parameters.

```{r, rq2, cache = TRUE}
## sigma in models
residual_sd <- 1

## effect size
min_bias_es <- es_outcome_smaller

## moderator beta
beta <- 0.11
## in our pilot data the beta (and simple correlation) between self-esteem and MGE was .11
## there are also larger estimates based on beta 0.419 from https://app.cooperationdatabank.org/
## YD: I can't get this link to load, what was this data, i.e. what is the DV if the IV is trust?
## this selected State Trust as independent variable, which as as of
## 2022-11-22, was composed of 11 effects and a total of 2036 participants

## we set the lowest value for the variance to 0.05

## for sensitivity analysis
## create list of random intercepts
list_random_intercepts <- c(0.05, 0.30, 0.55, 0.80, 1.5)

## select one value from list_rand_intercepts for example
random_intercept <- c(0.30)

## extract dataframe that has "n" participants across "c" countries
df <- fake_data_n200_c40

## fixed intercept 
fixed <- c(
    min_bias_es # outcome
  , beta # moderator
)

## random intercepts
random <- list(
    random_intercept  
)

## construct model
model_rq2 <- 
    
    makeLmer(
        min_bias ~ self_esteem + (1 | country), # we use simplier random structure to make code easier to run
        fixef = fixed, 
        VarCorr = random, 
        sigma = residual_sd, 
        data = df                 
    )

## check power
sim_rq2 <- 
    
    powerSim(
        model_rq2,
        test = fixed("self_esteem"),
        nsim = n_sim,
        alpha = 0.05/3
    )

sim_rq2

## extract number of countries and participants per country
n_country <- pull(count(unique(select(df, country))))
n_id <- pull(count(unique(select(df, id))))/n_country
            
## assign powersim to object for later use
summary_sim_rq2 <-
    
    summary(sim_rq2) %>%
    mutate(bias_es = min_bias_es,
           random = random_intercept,
           n_ids = n_id,
           n_countries = n_country
           )

## summarise
summary_sim_rq2
```

## Research question 3

We conduct a power analysis using the pre-specified mixed-effects model. For each fixed effect we specify the beta coefficients, for each random effect the variance, and finally the error variance (\ie sigma). The error variance is set to 1.

This is an example of estimating power for RQ3 for a given dataset using a single set of parameters.

Effects sizes relevant to national groups in economic tasks are a meta-analysis and a recent cross-cultural study. @Lane_2016 also eports an overall effect size of "national" groups $d = 0.16 (0.04–0.29; n = 52)$, and @Romano-et-al_2021 conducted a study across 42 nations finding $d = 0.22 (0.19-0.25)$ (this compares in-group with out-group plus strangers; from supplementary figures, it appears that the difference between in-group and out-group is less than in-group and stranger, so the overall effect size for in-group vs. out-group is likely less than $0.22$). We are not aware of an overall effect size of family from a meta-analysis or large scale study to report an effect size. Thus, we set the smaller effect
size to 0.16, which is the same used for RQ1 and RQ2.

```{r, rq3, cache = TRUE}
## sigma in models
residual_sd <- 1

## effect size
real_bias_es <- es_outcome_smaller

## extract dataframe that has "n" participants across "c" countries
df <-

    fake_data_n200_c40 %>%
    ## then adjust dataframe to contain relevant predictors (MGE, group type)
    pivot_longer(cols = c("nat_bias", "fam_bias"),
                 names_to = "group_type", values_to = "real_bias") %>%
    rename(MGE = min_bias) %>%
    separate(group_type, into = c("group_type", "bias")) %>%
    select(id, country, MGE, group_type, real_bias)

## moderators beta
beta_mge <- 0.13  # estimated strength of relationship between minimal group and real group bias,
# estimated from pilot data, lower of the two values where family was .13 and nationality was .36
beta_group_type <- 0.1 # difference between strength of national and family bias, estimate lower bound
beta_interaction <- 0.1 # differential relationship between MGE and bias for national vs family group, estimated smallest effect of interest

## we set the lowest value for the variance to 0.05 in our power
## analyss

## for sensitivity analysis
## create list of random intercepts
list_random_intercepts <- c(0.05, 0.30, 0.55, 0.80, 1.5)
## select one for example
random_intercept <- c(0.30)

## fixed intercept 
fixed <- c(
    real_bias_es # outcome
  , beta_mge # MGE
  , beta_group_type # group type
  , beta_interaction # interaction
)

## random intercepts
random <- list(
    random_intercept # for country
)

## construct model
## in final analyses, we will include
## (MGE + group_type | country):
## but we use a simplier specification for the power analysis
model_rq3 <- 
    
    makeLmer(
        real_bias ~ MGE * group_type +
            (1 | country), # we use simplier random structure to make code easier to run
        fixef = fixed, 
        VarCorr = random, 
        sigma = residual_sd, 
        data = df
    )

## check power for interaction
## YD: Really care most about power for MGE, how do we check that? 
## this is checking for power to detect the interaction...

sim_rq3 <- 
    
    powerSim(
        model_rq3,
        test =  fcompare(~ MGE + group_type),
        nsim = n_sim,
        alpha = 0.05/3
    )

sim_rq3




## extract number of countries and participants per country
n_country <- pull(count(unique(select(df, country))))
n_id <- pull(count(unique(select(df, id))))/n_country
            
## assign powersim to object for later use
summary_sim_rq3 <-
    
    summary(sim_rq3) %>%
    mutate(bias_es = real_bias_es,
           random = random_intercept,
           n_ids = n_id,
           n_countries = n_country
           )

## summarise
summary_sim_rq3
```

# References {-}

<div id="refs"></div>